A state space model for discovering the trajectory dynamics of replay

### 2020-03-11 Article.tex

\begin{figure*}%[tbhp]
\centering
\includegraphics[width=0.80\linewidth]{figures/Figure1/Figure1_v3}
\caption{Components of the state space model and demonstration on simulated data. \textbf{(A)} 19 simulated place cells with peaks evenly spaced on a 180 cm track that exhibit three different sequence patterns. For the first third of time, one cell fires repeatedly, representing one stationary location. For the second third of time, the cells fire in sequence, representing a smooth continuous trajectory through space. For the last third cells fire in an incoherent pattern, representing a non-smooth fragmented trajectory through space. \textbf{(B)}} Like the Bayesian Decoder, the state space encoding model uses estimates of the cell's place fields to compute a likelihood of position. \textbf{(C)} 
\label{Figure1}
\end{figure*}

\begin{figure*}%[tbhp]
\centering
\includegraphics[width=0.80\linewidth]{figures/Figure2/Figure2_v3}
\caption{Placeholder image of Iris with a long example caption to show justification setting.}
\label{Figure2}
\end{figure*}

\begin{figure*}%[tbhp]
\centering
\includegraphics[width=0.80\linewidth]{figures/Figure3/Figure3_v3}
\caption{Placeholder image of Iris with a long example caption to show justification setting.}
\label{Figure3}
\end{figure*}

\begin{figure*}%[tbhp]
\centering
\includegraphics[width=0.80\linewidth]{figures/Figure4/Figure4_v2}
\caption{Placeholder image of Iris with a long example caption to show justification setting.}
\label{Figure4}
\end{figure*}

>>>>

\begin{figure*}%[tbhp]
\centering
\includegraphics[width=0.80\linewidth]{figures/Figure1/Figure1_v3}
\caption{Components of the state space model and demonstration on simulated data. \textbf{(A)} 19 simulated place cells that exhibit three different sequence patterns. For the first third of time, one cell fires repeatedly, representing one stationary location. For the second third of time, the cells fire in sequence, representing a smooth continuous trajectory through space. For the last third cells fire in an incoherent pattern, representing a non-smooth fragmented trajectory through space. \textbf{(B)} Like the Bayesian Decoder, the state space model uses estimates of cells' place fields from when the animal is moving and combines them with the observed spikes in (A) to compute the likelihood of position for each time step. \textbf{(C)} This is combined with an explicit model of each dynamic state--which determines how latent position can change based on the posterior from the previous time step. In the hover state, this is modeled as the identity matrix because the position remains the same. In the continuous state, the dynamics are modeled as a Gaussian random walk, meaning the next possible position is likely to be spatially close. In the fragmented state, the dynamics are modeled as uniform, meaning the next position is equally likely to be anywhere. Note that the states displayed only reflect the within-state dynamics. \textbf{(D)} The probability of remaining in a particular state versus switching to another state. This is modeled as having a high probability of remaining in a particular state with a small probability of switching to one of the other states at each time step. \textbf{(F)} The state space model uses the components in A-D over all time to decode the joint probability of latent position and dynamic state. The joint probability can be summarized by marginalizing over latent position (left panel) to get the probability of each dynamic state over time (lines). This can also be used to get an estimate of latent position over time (right panel). \textbf{(G)} The probability of each state given a particular constant speed trajectory. Each speed corresponds to a simulated spiking sequence at that speed. The lines correspond to the average probability of that state over the sequence. Dotted line represents the 0.8 classification threshold we use to classify each range of speeds and shaded regions correspond to the classification we have given the range of speeds.
}
\label{Figure1}
\end{figure*}

\begin{figure*}%[tbhp]
\centering
\includegraphics[width=0.80\linewidth]{figures/Figure2/Figure2_v3}
\caption{Demonstration of the state space model on real data using both sorted and clusterless spikes on the same SWR. \textbf{(A)} Decoding using sorted spikes (arranged top to bottom). In the top panel, 31 cells on a W-track ordered according to linearized position by their place field peaks. The second panel shows the probability of each dynamic over time as in Figure 1F. Shaded regions correspond to the speed classifications in Figure 1G. The third panel shows the estimated probability of latent position over the course of the SWR as it travels down the center arm toward the right arm. L, R, C correspond to the position of the left, right and center wells respectively. The animal's position is indicated by the red triangle. \textbf{(B)} Decoding using clusterless spikes. The top panel shows multiunit spiking activity from each tetrode. Other panels have the same convention as(A).  \textbf{(C)} 1D MAP estimate of the latent position in (A) projected back into 2D. Color indicates time. The animal's position is denoted by the red circle. Grey lines show the animal's 2D position over the entire epoch. L, R, and C correspond to the left, right and center well as in (A). \textbf{(D)} 1D MAP estimate of the latent position in (B) projected back into 2D. Figure conventions are the same as in (C).
}
\label{Figure2}
\end{figure*}

\begin{figure*}%[tbhp]
\centering
\includegraphics[width=0.80\linewidth]{figures/Figure3/Figure3_v3}
\caption{A-D. Four examples of replays that would not be well-described using the Bayesian decoder. Figure conventions are the same as in Figure 2. \textbf{(A)} A replay that starts down the center arm away from the animal's position at the center well, slows down, and returns back. \textbf{(B)} A replay that stays at the left well for the entirety of the SWR while the animal is in the center arm. \textbf{(C)} A replay that begins stationary at the left well and then jumps to the middle of the right arm and proceeds up the right arm to the right well. \textbf{(D)} A replay that begins stationary at the left well, jumps to the center arm, proceeds away from the center well, jumps to the right arm, proceeds back toward the center well, and then becomes fragmented. \textbf{(E)} Percentage of SWRs in each epoch that has some content that falls into one of the categories (Hover, Hover-Continuous-Mix, Continuous, Fragmented-Continuous-Mix, or Fragmented). \textbf{(F)} Of the classified content, the percentage of SWRs in each epoch that have some Hover, Hover-Continuous-Mix, or Continuous content. \textbf{(G)} Of the classified content, the percentage of SWR in each epoch that has some Continuous content.
}
\label{Figure3}
\end{figure*}

\begin{figure*}%[tbhp]
\centering
\includegraphics[width=0.80\linewidth]{figures/Figure4/Figure4_v2}
\caption{\textbf{(A)} UpSet plot of the most common combinations of dynamics within each ripple. Each row represents a possible category present in the SWR. Each column represents a set of dynamics categories, where filled-in black dots with an edge between the dots indicates that multiple categories are present in the SWR (at temporally distinct times). The sets are ordered by how often they occurred as indicated by the bar plot above each category. The total number of each category is indicated by the rotated bar plot on the left. \textbf{(B)} Box plot of the duration of each dynamic category within a ripple. \textbf{(C)} The distance of the latent position from the animal's position for each dynamic category within each ripple. \textbf{(D)} Probability density estimate 
}
\label{Figure4}
\end{figure*}
